A comparative analysis of penalized linear mixed models in structured genetic data

### 2020-09-29 main.tex
Supplementary figures added and incorporated into fig refs

\subsubsection{Subpopulation structure}
For the fully simulated data, two subpopulation structures were considered, which we refer to as ``fine'' and ``coarse''.  We define a coarse structure to be one characterized by a few large subpopulations intended to model large-scale environmental heterogeneity and population stratification, such as that described in our motivating air quaility example. Conversely, we define a fine subpopulation structure to be one characterized by many small subpopulations, representing a continuum of environmental heterogeneity. Whether fine or coarse, the simulated data were generated such that subpopulations (also known as founding populations) are of equal size. We note that the distinction between fine and coarse structure is an entirely different concept than the level of admixture in the simulated data. It is the size of the subpopulations, not the level of admixture, that determines what we refer to as fine or coarse structure. For example, both the Independent Subpopulation and 1D Linear Admixture simulated data may be comprised of fine or coarse structure. However, the subpopulations will be more distinct in the unadmixed Independent Subpopulation data than in the 1D Linear Admixed data. The Empirical data is comprised of three racial subpopulations; one relatively large, one small, and one in the middle. Thus it represents a hybrid of our simulated coarse and fine structures, with little apparent admixture. Supplementary Figures \anna{?, ?, and ? MAKE THESE AND MAKE THEM PRETTY} illustrate the simulated kinship structures in the fine and coarse subpopulation settings, as well as that present in the empirical SNP data.

@@

LMM-Lasso-ggmix's poor performance in these scenarios are likely attributable to a number of a factors such as an ill-conditioned RRM  in $n \ge p$ settings \citep{ledoit2004well}. In addition, because LMM-Lasso-ggmix attempts to estimate more parameters than LMM-Lasso-Rakitsch, its optimization problem is not necessarily convex, making it prone to getting stuck in local minima before it is able to extensively explore the coefficient path. An examination of the number of iterations required for convergence for each $\lambda$ in the coefficient path suggests that this is the case for our various simulation settings (Supplementary Figure \ref{fig:s4}). Getting stuck in this manner would lead to few iterations required for each $\lambda$ value before convergence, allowing for an advantage in terms of computational speed, but poor estimation, as we see in Figure \ref{fig:eta_beta_mse}.

>>>>

\subsubsection{Subpopulation structure}
For the fully simulated data, two subpopulation structures were considered, which we refer to as ``fine'' and ``coarse''.  We define a coarse structure to be one characterized by a few large subpopulations intended to model large-scale environmental heterogeneity and population stratification, such as that described in our motivating air quaility example. Conversely, we define a fine subpopulation structure to be one characterized by many small subpopulations, representing a continuum of environmental heterogeneity. Whether fine or coarse, the simulated data were generated such that subpopulations (also known as founding populations) are of equal size. We note that the distinction between fine and coarse structure is an entirely different concept than the level of admixture in the simulated data. It is the size of the subpopulations, not the level of admixture, that determines what we refer to as fine or coarse structure. For example, both the Independent Subpopulation and 1D Linear Admixture simulated data may be comprised of fine or coarse structure. However, the subpopulations will be more distinct in the unadmixed Independent Subpopulation data than in the 1D Linear Admixed data. The Empirical data is comprised of three racial subpopulations; one relatively large, one small, and one in the middle. Thus it represents a hybrid of our simulated coarse and fine structures, with little apparent admixture. Supplementary Figures \ref{fig:admixed} - \ref{fig:empirical} illustrate the simulated kinship structures, as well as that for the Empirical SNP data.

@@

LMM-Lasso-ggmix's poor performance in these scenarios are likely attributable to a number of a factors such as an ill-conditioned RRM  in $n \ge p$ settings \citep{ledoit2004well}. In addition, because LMM-Lasso-ggmix attempts to estimate more parameters than LMM-Lasso-Rakitsch, its optimization problem is not necessarily convex, making it prone to getting stuck in local minima before it is able to extensively explore the coefficient path. An examination of the number of iterations required for convergence for each $\lambda$ in the coefficient path suggests that this is the case for our various simulation settings (Supplementary Figure \ref{fig:niter}). Getting stuck in this manner would lead to few iterations required for each $\lambda$ value before convergence, allowing for an advantage in terms of computational speed, but poor estimation, as we see in Figure \ref{fig:eta_beta_mse}.

### 2020-09-29 supplement.tex
Supplementary figures added and incorporated into fig refs

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Supplementary material}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[H]
    \centering
    \includegraphics[scale = 0.45]{figures/niter_perLambda_point.png}
    \caption{The number of iterations per $\lambda$ value for LMM-Lasso-Rakitsch and LMM-Lasso-ggmix across varying sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\eta = 0.8$, and $\xi = 0.8$, and dichotomous-discordant environmental effect structure. Note that 100 is the maximum number of iterations for both implementations of LMM-Lasso, so that numerous incidences of 100 iterations may indicate lack of model convergence.}
    \label{fig:s4}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Supplementary materials}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% 1D LINEAR ADMIXTURE KINSHIP
\begin{figure}[H]
    \centering
    \includegraphics[scale = 1]{figures/admixed_kinship.png}
    \caption{1D Linear Admixture with coarse subpopulation structure (50 subjects per subpopulation), estimated kinship matrix.}
    \label{fig:admixed}
\end{figure}

% INDEPENDENT SUBPOPULATIONS, COARSE, KINSHIP
\begin{figure}[H]
    \centering
    \includegraphics[scale = 1]{figures/indep_coarse_kinship.png}
    \caption{Independent Subpopulations with coarse subpopulation structure (50 subjects per subpopulation), estimated kinship matrix.}
    \label{fig:indep_coarse}
\end{figure}

% INDEPENDENT SUBPOPULATIONS, FINE, KINSHIP
\begin{figure}[H]
    \centering
    \includegraphics[scale = 1]{figures/indep_fine_kinship.png}
    \caption{Independent Subpopulations with subpopulation structure (2 subjects per subpopulation), estimated kinship matrix.}
    \label{fig:indep_fine}
\end{figure}

% EMPIRICAL KINSHIP
\begin{figure}[H]
    \centering
    \includegraphics[scale = 1]{figures/empirical_kinship.png}
    \caption{Empirical estimated kinship matrix.}
    \label{fig:empirical}
\end{figure}

\begin{landscape}
\begin{figure}[H]
    \centering
    \includegraphics[scale = 0.8]{figures/niter_perLambda_point.png}
    \caption{The number of iterations per $\lambda$ value for LMM-Lasso-Rakitsch and LMM-Lasso-ggmix across varying sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\eta = 0.8$, and $\xi = 0.8$, and dichotomous-discordant environmental effect structure. Note that 100 is the maximum number of iterations for both implementations of LMM-Lasso, so that numerous incidences of 100 iterations may indicate lack of model convergence.}
    \label{fig:niter}
\end{figure}
\end{landscape}
